\newpage
\TOCadd{Acknowledgements}

\chapter*{Acknowledgements}
\noindent I'd like to thank:

George Tzanetakis for suggesting masters research on synthesizers at UVic and for encouraging me to apply for an NSERC scholarship. I am thoroughly grateful for my time researching under your supervision and appreciate all the guidance I have received from you over the years.

Kirk McNally for encouraging me to apply for a JCURA during my last year of my undergrad and for supporting me to present work at my first international conferences. Those experiences opened me up to wanting to pursue graduate research. Thank you for all the support and guidance you've provided through my masters.

The Natural Sciences and Engineering Research Council of Canada, the British Columbia Provincial government, and the University of Victoria for the financial support provided to conduct this research.

My parents for their endless encouragement and support. Thank you Mom for encouraging me to pursue music and for helping me get Ableton during my first year of undergrad -- that is how this whole synthesizer business started. Thank you Dad for all the chats and support, and for helping with my writing during this thesis.

My loving partner Rachel for all the support and encouragement through my masters. Thank you for putting up with me when I was grumpy and for celebrating all the small victories with me along the way. And thank you for all the writing help.

Carson Gant for encouraging me to pursue music and for all the time you spent trying to guide me as I tried program that Flume sound into Massive synth. Much of this work is inspired by our time working on music together.

Joseph Turian, Max Henry, and Grace Brooks for your collaboration and support on the research presented as a part of this thesis. 

Colin Malloy and Keon Lee for being great lab partners. 

And thank you to everyone else I couldn't list. There are a lot of you!


% \noindent I would like to thank:
% \begin{description}
% \item[my cat, Star Trek, and the weather,]
% 	for supporting me in the low moments.
% \item[Supervisor Main,]
% 	for  mentoring, support, encouragement, and patience.
% \item[Grant Organization Name,]
% 	for funding me with a Scholarship.
% \end{description}

% \begin{flushright}
% \textit{I believe I know the only cure, which is to make
% one's centre of life inside of one's self, not
% selfishly or excludingly, but with a kind of
% unassailable serenity-to decorate one's inner house
% so richly that one is content there, glad to welcome
% any one who wants to come and stay, but happy all
% the same in the hours when one is inevitably alone.}
% \\
% Edith Wharton \\
% \end{flushright}
