\section{Forty Years of Automatic Synthesizer Programming}
- organized time of all the related work. Organized based on: synthesis type and method.
- Also include a separate table for evaluation methods (maybe not totally necessary).

\begin{itemize}
	\item \cite{zloof1977query} Early example of querying sounds (need to read)
	\item \cite{justice1979analytic} Coarse parameter matching by analyzing input audio - goal was to get in the ballpark and allow for parameter tweaking afterwards. FM (Single carrier with nested modulators). Hilbert Transform. Objective evaluation.
	\item \cite{wessel1979timbre} Not specifically for synthesis methods but classic paper. This should come much earlier in this section.
	\item \cite{beauchamp1982synthesis} Matching of alto sax/cornet sounds using an analytic method for FM synthesis. Looked at spectral centroid and RMS. Objective evaluation.
	\item \cite{ashley1986knowledge} A knowledge-based approach to assistance in timbral design (need to read still)
	\item \cite{payne1987microcomputer} Hilbert Transform (time domain). Also introduced FFT version with autocorrelation on spectrum. Periodic sampled sounds, FM DX7, objective evaluation.
	\item \cite{delprat1990parameter} Parameter estimation for non-linear resynthesis methods with the help of a time-frequency analysis of natural sounds.
	\item \cite{horner1993machine} Genetic algorithms for inverse synthesis of instrumental sounds on FM synthesis engine.
	\item \cite{horner1993methods} Wavetable
	\item \cite{vuori1993parameter} Parameter estimation of non-linear physical models by simulated evolution-application to the flute model
	\item \cite{takala1993using} Using Physically-Based Models and Genetic Algorithms for Functional Composition of Sound Signals
	\item \cite{fujinaga1994genetic} Genetic algorithms as a method for granular synthesis regulation
	\item \cite{ethington1994seawave} Semantic search -- this is an good one, as far as I can tell so far this is the first semantic search for synthesizer sounds.
	\item \cite{miranda1995artificial} An artificial intelligence approach to sound design
	\item \cite{horner1995wavetable} Updated version of wavetable matching. Used multiple pitches of instrumental tones and a genetic algorithm.
	\item \cite{horner1995envelope} Envelope matching with genetic algorithms
	\item \cite{horner1996computation} Computation and memory tradeoffs with multiple wavetable interpolation
	\item \cite{horner1996piecewise} Piecewise-linear approximation of additive synthesis envelopes: a comparison of various methods
	\item \cite{cheung1996group} Group synthesis (wavetable) with genetic algorithms
	\item \cite{horner1996double} Double-modulator FM matching of instrument tones
\end{itemize}


\subsection{Synthesis Type}
An overview of synthesis type that was the focus of automatic synthesizer programming studies.

\subsubsection{FM}
\cite{justice1979analytic}\cite{beauchamp1982synthesis}\cite{payne1987microcomputer}\cite{horner1993machine}\cite{horner1996double}\cite{tan1996automated}\cite{delprat1997global}\cite{lim1999performance}\cite{tan2003automated}
\cite{mitchell2005frequency}\cite{mitchell2007evolutionary}\cite{clement2011automatic}\cite{roth2011comparison}\cite{macret2012automatic}\cite{hamadicharef2012intelligent}\cite{barkan2019deep}

\subsubsection{Non-linear}
\cite{beauchamp1982synthesis}\cite{delprat1990parameter}

\subsubsection{Wavetable / Group Synthesis}
\cite{horner1993methods}\cite{horner1995wavetable}\cite{horner1995envelope}\cite{horner1996computation}\cite{horner1996piecewise}\cite{cheung1996group}\cite{oates1997analytical}\cite{horner1998modeling}\cite{lee1999modeling}\cite{so2002wavetable}

\subsubsection{Physical Modelling}
\cite{vuori1993parameter}\cite{erkut2000extraction}\cite{liang2000recurrent}\cite{nackaerts2001parameter}\cite{riionheimo2003parameter}

\subsubsection{Granular}
\cite{fujinaga1994genetic}\cite{johnson1999exploring}

\subsubsection{Additive}
\cite{ethington1994seawave}\cite{horner1995envelope}\cite{horner1996piecewise}\cite{johnson2006timbre}\cite{mintz2007toward}

\subsubsection{Subtractive}
\cite{roth2011comparison}

\subsubsection{Generic VST}
\cite{yee2008synthbot}\cite{heise2009automatic}

\subsubsection{Other}
Noise-band \cite{chinen2007genesynth}
Concatenative \cite{stowell2010making}
Teenage Engineering OP-1 (Multiple Synthesis Engines) \cite{macret2013automatic}

\subsection{Estimation Method}
An overview of the method used to select a synthesizer parameter setting based on input.

\subsubsection{Analytic / Signal Processing}
\cite{justice1979analytic}\cite{beauchamp1982synthesis}\cite{payne1987microcomputer}\cite{ethington1994seawave}

\subsubsection{Genetic}
\cite{horner1993machine}\cite{fujinaga1994genetic}\cite{horner1995envelope}\cite{horner1995wavetable}\cite{riionheimo2003parameter}\cite{mandelis2003musical}\cite{mitchell2005frequency}\cite{mitchell2007evolutionary}
\cite{chinen2007genesynth}\cite{yee2008synthbot}\cite{roth2011comparison}\cite{macret2012automatic}\cite{hamadicharef2012intelligent}\cite{macret2013automatic}

\subsubsection{Interactive Genetic}
\cite{johnson1999exploring}

\subsubsection{Neural Network}
\cite{johnson2006timbre}\cite{roth2011comparison}\cite{zhang2018visualization}\cite{barkan2019deep}

\subsubsection{Data-driven}
\cite{roth2011comparison}\cite{mcartwright2014}

\subsubsection{Other}
Linear coding \cite{mintz2007toward}
Particle Swarm Optimization \cite{heise2009automatic}\cite{munoz2011opposition}
Regression Tree \cite{stowell2010making}
Semantic Clustering \cite{clement2011automatic}
Hill-Climber \cite{roth2011comparison}

\subsection{Unsorted}
These references have not been reviewed or sorted yet!

\cite{takala1993using}
\cite{hourdin1997sound}
\cite{horner1998nested}
\cite{wehn1998using}
\cite{garcia2001automatic}
\cite{dahlstedt2001creating}
\cite{garcia2001growing}
\cite{jehan2001perceptual}
\cite{su2002class}
\cite{le2002neural}
\cite{arfib2002strategies}
\cite{miranda2004crossroads}
\cite{schatter2005synaesthetic}
\cite{gounaropoulos2006synthesising}
\cite{lai2006automated}
\cite{mcdermott2007evolutionary}
\cite{yee2007automated}
\cite{yee2007evolving}
\cite{howard2007timbral}
\cite{mcdermott2008evolutionary}
\cite{yee2011automatic}
\cite{mitchell2012automated}
\cite{povscic2013controlling}
\cite{seago2013new}
\cite{krekovic2014intelligent}
\cite{macret2014automatic}
\cite{itoyama2014parameter}
\cite{huang2014active}
\cite{fasciani2016tsam}
\cite{krekovic2016algorithm}
\cite{tatar2016automatic}
\cite{yee2016use}
\cite{smith2017play}
\cite{yee2018automatic}
\cite{luke2019stochastic}
